---
title: "Salmon Biphasic model 4"
author: "Viktor Thunell" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    page-layout: full
    embed-resources: true
    toc: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

## Load libraries

```{r libs}
#| message: false
#| warning: false
#| cache: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog", "devtools","viridis","nls.multstart", "broom", "patchwork", "coda", "boot", "tidybayes","bayesplot", "nimble", "here")

# remotes::install_github("nimble-dev/nimble", ref = "devel", subdir = "packages/nimble")

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){
    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  }

invisible(lapply(pkgs, library, character.only = T))

options(ggplot2.continuous.colour = "viridis")
theme_set(theme_light()) # check why global theme option not working when rendering
# Set path
color_scheme_set("viridis")
home <- here::here()
```

## Read data

```{r}
#| message: false
#| warning: false
#| cache: false

sallaa <- readRDS(file = paste0(home,"/data/data-for-2-2/salmon-laa_2025-09-08.RData")) %>%
  rename(year = fi_year,
         site = sai_location,
         sea.age = sea_age_year,
         juv.age = juvenile_age_year,
         tot.age = tot_age_year,
         lat = fisa_y_4326,
         lon = fisa_x_4326) %>%
  #filter out non-aged individuals (~90000 individuals)
  filter(!(is.na(sea.age) & is.na(juv.age))) %>%
  filter(!is.na(juv.age),
         !is.na(year),
         year > 2000
         ) %>%
  slice_sample(n = 20000)

```

## Plot data
```{r plot dat}
sallaa %>%
  drop_na(sex) %>%
  ggplot(aes(tot.age, length_mm, color = sex)) +
  geom_point() +
  facet_grid(age.type~sex) +
  scale_x_continuous(breaks = seq(0, 13, 1) ) +
  theme_light() 

# If dropping all inds where both sea and sm age is NA
sallaa %>% 
  ggplot() +
  geom_density(aes(x = tot.age), fill = "deeppink", alpha = 0.5) +
  xlim(0, 13) +

sallaa %>% 
  ggplot() + 
  geom_density(aes(x = length_mm), fill = "deeppink", alpha = 0.5)

sallaa %>% 
  drop_na(sea.age) %>%
  ggplot() +
  geom_density(aes(x = length_mm, fill = factor(tot.age)), alpha = 0.5) +
  facet_wrap(~tot.age)

# age at smoltification by spatial unit
sallaa %>% 
  filter(age.type == "both") %>%
  ggplot() +
  geom_density(aes(x = juv.age, color = spat.unit),  alpha = 0.5) +
  facet_wrap(~spat.unit) +
  xlim(-1, 5)

# juvenile size at age
sallaa %>%
  filter(age.type == "juve.only") %>%
  ggplot(aes(x = juv.age, y =length_mm, color = spat.unit)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm) +
  facet_wrap(~spat.unit)

# size at age zero
sallaa %>%
  filter(!age.type == "sea.only",
         tot.age == 0) %>%
  ggplot() +
  geom_density(aes(x = length_mm), fill = "deeppink", alpha = 0.5) +
  labs(title = "Age 0")
  
```


## Check decimal age effect on length

```{r}
# The majority of age 0 fish are caught in autumn
sallaa %>%
  filter(!is.na(tot.age),
         tot.age == 0) %>%
  mutate(yday = yday(date)) %>%
  count(yday) %>%   # counts per day
  ggplot(aes(x = yday, y = n, group = 1)) +
  geom_line(color = "steelblue") +
  geom_vline(xintercept = 365/2, color = "red")

sallaa %>%
  filter(!is.na(tot.age)) %>%
  ggplot() +
  geom_density(aes(x = tot_age_decimyear), fill = "deeppink", alpha = 0.5) +
  facet_wrap(~tot.age, scales = "free_y")

# But there seems to be a variable effect of time of the year on length
sallaa %>%
  filter(age.type == "both") %>%
  ggplot(aes(x = tot_age_decimyear, y = length_mm, color = spat.unit)) +
  #geom_point(alpha = 0.5) +
  facet_wrap(~tot.age, scales = "free_y") +
  stat_smooth(method = "lm", se = FALSE) #+
  #stat_smooth(method = "gam", formula = y ~ s(x, k=3))
  ylim(c(400,1200))

sallaa %>%
  filter(age.type == "juve.only") %>%
  ggplot(aes(x = tot_age_decimyear, y = length_mm, color = spat.unit)) +
  #geom_point(alpha = 0.5) +
  facet_wrap(~tot.age, scales = "free_y") +
  stat_smooth(method = "lm", formula = y ~ x, se = FALSE)

# all spatial units grouped
sallaa %>%
  filter(age.type == "both") %>%
  ggplot(aes(x = tot_age_decimyear, y = length_mm, color = factor(tot.age))) +
  stat_smooth(method = "lm",se = FALSE) +
  #stat_smooth(method = "gam", formula = y ~ s(x, k=3)) +
  labs(title = "length regressed over time of year by age class")
```

## Biphasic 5a - without plus growth 

Assuming individuals are born May 1.
Spatiotemporal model for g and par

```{r}

biph5.code <- nimbleCode({

  # likelihood
  for(i in 1:nobs){
    length_mm[i] ~ dnorm(l.mu[age.index[i], ind.id[i]], sd = sig.l)
    }

  # Estimate length at age for each individual (j in n.ind) and for each age of j (k in age.index=1 to age.index[j], i.e. age 0 to catch age)
  for(i in 1:n.ind){
    
    l.mu[1,i] <- lb.mu[spat.unit[i]] + exp(g[spat.unit[i], hatch.year.f[i]]) # age 0
    
    #length prediction 1+ and older
    for(a in 2:(age.index[i]+1)){   #age.index should be obs age +1
      
      l.mu[a,i] <- l.mu[a-1,i] + step(smo.age[i] - (a-1))*exp(g[spat.unit[i], (hatch.year.f[i]+(a-1))]) +
        (1-step(smo.age[i] - (a-1)))*((exp(par[spat.unit[i],(hatch.year.f[i]+(a-1)),1]) - l.mu[(a-1),i])*(1 - exp(-exp(par[spat.unit[i],(hatch.year.f[i]+(a-1)),2]))))
      
      }
    }
  
  # estimate smo.age
  for(l in 1:n.ind){
    smo.age[l] ~ dcat(smp[1:nK,spat.unit[l]])
    #smo.age[l] ~ dpois(lam.sm[spat.unit[l]]) # old
    }

  # Priors
  sig.l ~ dexp(1/100)
  
  # LKJ prior on correlation matrix, see NIMBLE manual p45.
  Ustar[1:npars,1:npars] ~ dlkj_corr_cholesky(1.3, npars) # eta = 1.3
  U[1:npars,1:npars] <- uppertri_mult_diag(Ustar[1:npars, 1:npars], sig.par[1:npars])

  # priors for smolt age and juvenile growth rate
  #sc.sm <- 0.5^2/2.5 # shape = 2.5^2/.5^2 = 25, #From mean=shape*scale, var=shape*scale^2 -> shape = mean^2/var & scale = var/mean. mean=2.5, sd=0.5.

  for(i in 1:nsu){
    smp[1:nK,i] ~ ddirch(alpha[1:nK]) 
    lb.mu[i] ~ dnorm(17, sd = 5) # ~15 to 20 mm, https://doi.org/10.1111/j.1095-8649.2009.02497.x sd is guesstimate
    
    for(j in minyear:maxyear){
      par[i,j,1:npars] ~ dmnorm(mu.par[1:npars], cholesky = U[1:npars, 1:npars], prec_param = 0)
      g[i,j] ~ dnorm(mean = g.lmean, sd = g.lsd) 
    }
  }

  g.lsd <- sqrt(log(1 + (10^2) / (50^2))) # sqrt of variance (= sd) of the lognormal distr. 50 and 10 is arbitrary atm.
  g.lmean <- log(50) - 0.5 * g.lsd^2  # mean of the lognorm distr.
  
  mu.par[1] ~ dnorm(mean = l.lmean, sd = l.lsd)
  mu.par[2] ~ dnorm(mean = k.lmean, sd = k.lsd)
  sig.par[1] ~ dlnorm(0,1)
  sig.par[2] ~ dlnorm(0,1)

  # K & l values from fb  
  l.lsd <- sqrt(log(1 + (278^2) / (1378^2)))
  l.lmean <- log(1378) - 0.5 * l.lsd^2
  k.lsd <- sqrt(log(1 + (0.28^2) / (0.43^2)))
  k.lmean <- log(0.43) - 0.5 * k.lsd^2  
  })

# Function creating the Cholesky of the covar. matrix (p45 Nimble manual)
uppertri_mult_diag <- nimbleFunction(
  run = function(mat = double(2), vec = double(1)) {
    returnType(double(2))
    p <- length(vec)
    out <- matrix(nrow = p, ncol = p, init = FALSE)
    for(k in 1:p)
      out[ , k] <- mat[ , k] * vec[k]
    return(out)
   # turn off buildDerivs for the i index
}, buildDerivs = list(run = list(ignore = c('k'))))

data <- sallaa %>%
  mutate(smo.age = if_else(age.type == "both", juv.age, NA),
         sea.age = if_else(is.na(sea.age), 0, sea.age),
         age.index = as.integer(tot.age + 1), # index for l.mu, + 1 to include age 0 individuals, add 0.1 to test "tot.age %% 1"
         hatch.year = year-tot.age,
         hatch.year.f = as.numeric(factor(hatch.year, levels = sort(unique(hatch.year)), labels = seq_along(sort(unique(hatch.year))))),
         year.f = as.integer(factor(year)),
         spat.unit = as.integer(factor(spat.unit))) %>%
  select(smo.age, tot.age, length_mm, age.index, year, spat.unit, hatch.year, hatch.year.f) %>%
  mutate(ind.id = row_number()) # ind id

n.ind <- nrow(data) # n individuals
npars <- 2 # n parameters in the LKJ dmnorm 
nsu <- length(unique(data$spat.unit))
age.index = data$age.index
n.ages = length(unique(age.index))+1
nK = length(unique(data$smo.age %>% na.omit()))
# get max year by, needed as hatch.year is integer and the last year is not (simpler solution?).
maxyear <- max(data$hatch.year.f) + data %>%
  filter(year == max(year)) %>%
  filter(tot.age == max(tot.age)) %>%
  distinct(tot.age) %>% pull(tot.age)
minyear = min(data$hatch.year.f)

# biph5.model$initializeInfo()
inits <- function(){
  list(#l.mu = matrix(abs(rnorm(n.ages * n.ind,500,25)), nrow = n.ages, ncol = n.ind), # this is not correct, only inits of =<tot.age
       sig.l = abs(rnorm(1,250,25)),
       sig.par = c(rlnorm(1,0,1),
                   rlnorm(1,0,1)),
       mu.par = c(rnorm(1,7,0.2),
                  rnorm(1,-1,0.6)),
       lb.mu = abs(rnorm(nsu,20,5)),
       g = matrix(rnorm(nsu * length(minyear:maxyear),4,.2), nrow = nsu, ncol = length(minyear:maxyear)),
       smo.age = round(runif(n.ind,0,nK))
       #Ustar, #U, #par
       )
  }
       
# build model
biph5.model <- nimbleModel(biph5.code,
                           constants = list(npars = npars,
                                            nobs = nrow(data),
                                            n.ind = n.ind, # change when >1 sample per ind.id
                                            nsu = nsu,
                                            nK = nK,
                                            alpha = c(1, 2, 3, 2, 1, 1)[1:nK],
                                            age.index = age.index,
                                            ind.id = data$ind.id,
                                            hatch.year.f = data$hatch.year.f,
                                            minyear = minyear,
                                            maxyear = maxyear,
                                            spat.unit = data$spat.unit),
                           inits = inits(),
                           data = data %>% select(-ind.id,-age.index,-spat.unit,-hatch.year,-tot.age, -year, -hatch.year.f))

# check values
biph5.model$initializeInfo()
biph5.model$simulate("sig.l")
biph5.model$simulate("mu.par")
biph5.model$simulate("sig.par")
biph5.model$simulate("Ustar")
biph5.model$simulate("U")
biph5.model$simulate("smp")
biph5.model$simulate("smo.age")
biph5.model$simulate("lb.mu")
biph5.model$simulate("g")
biph5.model$simulate("par")
biph5.model$simulate("l.mu")

# biph5.model$calculate("par[,,1]")
# biph5.model$calculate(biph5.model$getDependencies(c("par[1,2,1]")))
# biph5.model$calculate(par)
# biph5.model$calculate("sig.l")
# biph5.model$calculate("mu.par[1]")
# biph5.model$calculate("sig.par")
# biph5.model$calculate("Ustar")
# biph5.model$calculate("U")
# biph5.model$calculate("smp")
# biph5.model$calculate("smo.age")
# biph5.model$calculate("lb.mu")
# biph5.model$calculate("g")
# biph5.model$calculate(par[1:10,1,1])
# biph5.model$calculate("l.mu")

biph5.model$sig.l
biph5.model$U
biph5.model$Ustar
biph5.model$mu.par
biph5.model$sig.par
biph5.model$smp
biph5.model$smo.age
biph5.model$lb.mu
biph5.model$par
exp(biph5.model$g)
exp(biph5.model$par[1:11,1:25,1])
exp(biph5.model$par[1:11,1:20,2])
# Negative l.mus
sum(is.na(biph5.model$l.mu))/length(biph5.model$l.mu) # percent negatives
# NAs are after
biph5.model$l.mu[,1000:1100] # 

```

<!-- ### Non PROJECT Sample g, par, lb.mu, sig.l -->

<!-- ```{r samp 1} -->
<!-- # configure.. -->
<!-- biph5.confmcmc <- configureMCMC(biph5.model, monitors = c("g","par","lb.mu","sig.l")) -->

<!-- # and build MCMC -->
<!-- biph5.mcmc <- buildMCMC(biph5.confmcmc) -->

<!-- # compile model -->
<!-- biph5.c <- compileNimble(biph5.model) -->

<!-- # compile mcmc (and specify the project model) -->
<!-- biph5.mcmcc <- compileNimble(biph5.mcmc) -->

<!-- # MCMC Samples -->
<!-- #biph5.samp <- runMCMC(biph5.mcmcc, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE) -->
<!-- biph5.samp <- runMCMC(biph5.mcmcc, niter = 3000, nburnin = 1000, nchains = 2, samplesAsCodaMCMC = TRUE) -->

<!-- ``` -->
### Sample g, par, lb.mu, sig.l

```{r samp 1}
# compile model
biph5.c <- compileNimble(biph5.model)

# configure..
biph5.confmcmc <- configureMCMC(biph5.c, monitors = c("g","par","lb.mu","sig.l"), useConjugacy = FALSE, project = biph5.model, enableWAIC = TRUE)

# and build MCMC
biph5.mcmc <- buildMCMC(biph5.confmcmc, project = biph5.model)

# compile mcmc
biph5.mcmcc <- compileNimble(biph5.mcmc, project = biph5.model)

# MCMC Samples
#biph5.samp <- runMCMC(biph5.mcmcc, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE)
biph5.samp <- runMCMC(biph5.mcmcc, niter = 6000, nburnin = 3000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)

```

### Sample l.mu
```{r samp 2}
# # configure...
# biph5.confmcmc <- configureMCMC(biph5.c, monitors = c("l.mu"), project = biph5.model, enableWAIC = TRUE)
# 
# # and build MCMC
# biph5.mcmc.lmu <- buildMCMC(biph5.confmcmc, project = biph5.model)
# 
# # compile mcmc (and specify the project model)
# biph5.mcmcc.lmu <- compileNimble(biph5.mcmc.lmu, project = biph5.model, resetFunctions = TRUE)
# 
# # MCMC Samples
# #biph5.samp.lmu <- runMCMC(biph5.mcmcc.lmu, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE)
# biph5.samp.lmu <- runMCMC(biph5.mcmcc.lmu, niter = 6000, nburnin = 3000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)

biph5.c <- compileNimble(biph5.model)
biph5.confmcmc <- configureMCMC(biph5.model, monitors = c("l.mu"), useConjugacy = FALSE, enableWAIC = TRUE)

# and build MCMC
biph5.mcmc.lmu <- buildMCMC(biph5.confmcmc)

# compile mcmc (and specify the project model)
biph5.mcmcc.lmu <- compileNimble(biph5.mcmc.lmu)

# MCMC Samples
#biph5.samp.lmu <- runMCMC(biph5.mcmcc.lmu, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE)
biph5.samp.lmu <- runMCMC(biph5.mcmcc.lmu, niter = 5000, nburnin = 3000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)
```


### Sample smo.age
```{r samp 3}
# configure..
biph5.confmcmc <- configureMCMC(biph5.c, monitors = c("smo.age"), project = biph5.model, enableWAIC = TRUE)

# and build MCMC
biph5.mcmc.smo <- buildMCMC(biph5.confmcmc, project = biph5.model)

# compile mcmc (and specify the project model)
biph5.mcmcc.smo <- compileNimble(biph5.mcmc.smo, project = biph5.model)

# MCMC Samples
#biph5.samp.lmu <- runMCMC(biph5.mcmcc.lmu, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE)
biph5.samp.smo <- runMCMC(biph5.mcmcc.smo, niter = 6000, nburnin = 3000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)
```

### Check g, par, lb.mu, sig.l
```{r check pars}
samp <- biph5.samp$samples
sum <- summary(samp)

# summary of values of sampled parameters
vars_subset <- grep("^g\\[", rownames(sum$statistics), value = TRUE)
summary(exp(sum$statistics[which(rownames(sum$statistics)%in%vars_subset),][,1]))
vars_subset <- grep("^par\\[", rownames(sum$statistics), value = TRUE)
vars_subset1 <- grep("1]", vars_subset, value = TRUE)
summary(exp(sum$statistics[which(rownames(sum$statistics)%in%vars_subset1),][,1]))
vars_subset2 <- grep("2]", vars_subset, value = TRUE)
summary(exp(sum$statistics[which(rownames(sum$statistics)%in%vars_subset2),][,1]))
vars_subset <- grep("^lb.mu\\[", rownames(sum$statistics), value = TRUE)
summary(sum$statistics[which(rownames(sum$statistics)%in%vars_subset),][,1])
vars_subset <- grep("sig.l", rownames(sum$statistics), value = TRUE)
sum$statistics[which(rownames(sum$statistics)%in%vars_subset),]

# some trace plots
rn <- sample(1:200, 10)
vars_subset <- grep("^g\\[", varnames(samp), value = TRUE)[rn]
traceplot(samp[, vars_subset, drop = FALSE])

vars_subset <- grep("^par\\[", varnames(samp), value = TRUE)[rn]
traceplot(samp[, vars_subset, drop = FALSE])

vars_subset <- grep("^lb.mu\\[", varnames(samp), value = TRUE)
traceplot(samp[, vars_subset, drop = FALSE])

vars_subset <- grep("sig.l", varnames(samp), value = TRUE)
traceplot(samp[, vars_subset, drop = FALSE])

```

### Plot g, par...

```{r plot pars}
# g[spat.unit, hatch.year]
samp %>%
  gather_draws(g[su, hy], sep = ",") %>%
  filter(hy == c(1:5)) %>%
  ggplot() +
  geom_density(aes(x = exp(.value), color = factor(hy))) + # age parameter
  facet_wrap(~su, scales = "free") +
  theme_light()

# lb.mu[spat.unit]
samp %>%
  gather_draws(lb.mu[su], sep = ",") %>%
  ggplot() +
  geom_density(aes(x = .value, color = factor(su))) + # age parameter
  theme_light()

# par
samp %>%
  gather_draws(par[su, hy, parm], sep = ",") %>%
  filter(hy == c(1:5)) %>%
  ggplot() +
  geom_density(aes(x = exp(.value), color = factor(hy))) + # age parameter
  facet_wrap(parm~su, scales = "free") +
  theme_light() +
  labs("linf and K")

# # growth curves not looking ok. Should plot the individual curves and summarise the mean by su curve
# samp %>%
#   spread_draws(par[su,hy,p], g[su,hy], lb.mu[su], sep = ",") %>%
#   mutate(g = exp(g),
#          par = exp(par)) %>%
#   group_by(.draw,su,p) %>%
#   summarise(par.m = mean(par), g.m = mean(g), lb.mu = mean(lb.mu), .groups = "keep") %>%
#   group_by(su,p) %>%
#   median_qi() %>%
#   mutate(p = case_when(p == 1 ~ "Linf",
#                        p == 2 ~ "K",
#                        .default = NA)) %>%
#   pivot_wider(id_cols = su, names_from = p, values_from = c(par.m, g.m, lb.mu)) %>%
#   expand_grid(age = seq(0, 12, 0.1)) %>%
#   mutate(length.a = par.m_Linf*(1-exp(-par.m_K*age)),
#          length.j = lb.mu_Linf + g.m_Linf) %>%
#   ggplot() +
#   geom_point(data = sallaa %>% filter(!age.type == "sea.only"), 
#              aes(tot.age, length_mm), alpha = 0.1, color = "lightblue") +
#   geom_line(aes(age, length.a, color = factor(su))) +
#   geom_line(aes(age, length.j, color = factor(su))) +
#   theme_light() 

```

### Check l.mu
```{r check lmu}
#samp.lmu <- biph5.samp.lmu
# remove NA nodes (the older-than-catch-age nodes)
cm <- lapply(biph5.samp.lmu$samples, as.matrix)
nonNA_nodes <- cm %>% 
  map(~ colnames(.x)[colSums(is.na(.x)) == 0]) %>%
  reduce(intersect) # applies function intersect over elements of the lists returned by map(), i.e. removing the colnames from both lists.
cm_cleaned <- lapply(cm, function(x) x[, nonNA_nodes, drop = FALSE])
rm(cm) # big one
samp.cl <- mcmc.list(lapply(cm_cleaned, mcmc))  

#sum <- summary(samp.cl) # not working when l.mu == NA

# some traces of l.mu
vars_subset <- grep("^l.mu\\[", varnames(samp.cl), value = TRUE)[rn]
traceplot(samp.cl[, vars_subset, drop = FALSE])

```

### Plot predicted vs observed l.mu

```{r predobs}
# plot observed against predicted
obs <- data %>%
  select(length_mm,tot.age,ind.id,spat.unit) %>%
  mutate(tot.age = round(tot.age),
         observed = length_mm)
gc()
predi <- samp.cl[,1:2000] %>%
  gather_draws(l.mu[age.i,ind.id], sep = ",") %>%
  median_qi() %>%
  rename(predicted = .value) %>%
  mutate(tot.age = age.i-1) %>%
  filter(tot.age %in% obs$tot.age & ind.id %in% obs$ind.id)

inte.po <-
predi %>%
  left_join(obs, by = join_by(tot.age, ind.id)) %>%
  ggplot(aes(x = observed, y = predicted)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_errorbar(aes(ymin = .lower, ymax = .upper), width = 0.2, alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_minimal() +
  ylim(0,1300) +
  xlim(0,1300)


predi %>%
  left_join(obs, by = join_by(tot.age, ind.id)) %>%
  group_by(ind.id) %>%
  mutate(spat.unit = ifelse(is.na(spat.unit), first(na.omit(spat.unit)), spat.unit)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(tot.age, predicted, color = ind.id, group = ind.id)) +
  geom_point(aes(tot.age, observed, color = ind.id, group = ind.id), alpha = 0.2) +
  facet_wrap(~spat.unit)
```

### Plot smolt ages
```{r }
data %>%
  filter(!is.na(smo.age)) %>%
  select(smo.age, spat.unit, ind.id) %>%
  mutate(t= "obs") %>%
  bind_rows( biph5.samp.smo %>% 
               gather_draws(smo.age[ind.id]) %>% 
               rename(smo.age = .value) %>% 
               mutate(t="pred") %>%
               left_join(data %>% select(spat.unit, ind.id), by = "ind.id") ) %>%
  ggplot() +
  geom_density(aes(smo.age, fill = t)) +
  facet_wrap(~spat.unit)
```

## Biphasic 5b - with plus growth 

Assuming individuals are born May 1.
Spatiotemporal model for g and par

```{r}

biph5.con.code <- nimbleCode({

  # likelihood
  for(i in 1:nobs){
    length_mm[i] ~ dnorm(l.mu[age.index[i], ind.id[i]], sd = sig.l)
    }

  # Estimate length at age
  for(i in 1:n.ind){
    #length prediction 0+
    l.mu[1,i] <- lb.mu[spat.unit[i]] + 
      step(1-tot.age.2[i])*exp(g[spat.unit[i], hatch.year.f[i]])*doy.dec[i] +
      (1-step(1-tot.age.2[i]))*exp(g[spat.unit[i], hatch.year.f[i]])

    #length prediction 1+ and older
    for(a in 2:(age.index[i]+1)){
      l.mu[a,i] <- l.mu[(a-1),i] +
        
        step(a-tot.age.2[i])*(step(smo.age[i] - (a-1))*exp(g[spat.unit[i], (hatch.year.f[i]+(a-1))])*doy.dec[i] + (1-step(smo.age[i] - (a-1)))*((exp(par[spat.unit[i],(hatch.year.f[i]+(a-1)),1]) - l.mu[(a-1),i])*(1 - exp(-exp(par[spat.unit[i],(hatch.year.f[i]+(a-1)),2])*doy.dec[i])))) +

        (1-step(a-tot.age.2[i]))*(step(smo.age[i] - (a-1))*exp(g[spat.unit[i], (hatch.year.f[i]+(a-1))]) + (1-step(smo.age[i] - (a-1)))*((exp(par[spat.unit[i],(hatch.year.f[i]+(a-1)),1]) - l.mu[(a-1),i])*(1 - exp(-exp(par[spat.unit[i],(hatch.year.f[i]+(a-1)),2])))))

          }
    }
  
  # estimate smo.age
  for(l in 1:n.ind){
    smo.age[l] ~ dcat(smp[1:nK,spat.unit[l]])
    }

  # Priors
  sig.l ~ dexp(1/100)
  
  # LKJ prior on correlation matrix, see NIMBLE manual p45.
  Ustar[1:npars,1:npars] ~ dlkj_corr_cholesky(1.3, npars) # eta = 1.3
  U[1:npars,1:npars] <- uppertri_mult_diag(Ustar[1:npars, 1:npars], sig.par[1:npars])

  # priors for smolt age and juvenile growth rate
  for(i in 1:nsu){
    smp[1:nK,i] ~ ddirch(alpha[1:nK])  
    lb.mu[i] ~ dnorm(17, sd = 5) # ~15 to 20 mm, https://doi.org/10.1111/j.1095-8649.2009.02497.x sd is guesstimate
    for(j in minyear:maxyear){
      par[i,j,1:npars] ~ dmnorm(mu.par[1:npars], cholesky = U[1:npars, 1:npars], prec_param = 0)
      g[i,j] ~ dnorm(mean = g.lmean, sd = g.lsd) 
    }
  }

  g.lsd <- sqrt(log(1 + (10^2) / (50^2))) # sqrt of variance (= sd) of the lognormal distr. 50 and 10 is arbitrary atm.
  g.lmean <- log(50) - 0.5 * g.lsd^2  # mean of the lognorm distr.
  
  mu.par[1] ~ dnorm(mean = l.lmean, sd = l.lsd)
  mu.par[2] ~ dnorm(mean = k.lmean, sd = k.lsd)
  sig.par[1] ~ dlnorm(0,1)
  sig.par[2] ~ dlnorm(0,1)

  # K & l values from fb  
  l.lsd <- sqrt(log(1 + (278^2) / (1378^2)))
  l.lmean <- log(1378) - 0.5 * l.lsd^2
  k.lsd <- sqrt(log(1 + (0.28^2) / (0.43^2)))
  k.lmean <- log(0.43) - 0.5 * k.lsd^2  
  
  })

# Function creating the Cholesky of the covar. matrix (p45 Nimble manual)
uppertri_mult_diag <- nimbleFunction(
  run = function(mat = double(2), vec = double(1)) {
    returnType(double(2))
    p <- length(vec)
    out <- matrix(nrow = p, ncol = p, init = FALSE)
    for(k in 1:p)
      out[ , k] <- mat[ , k] * vec[k]
    return(out)
   # turn off buildDerivs for the i index
}, buildDerivs = list(run = list(ignore = c('k'))))


data <- sallaa %>%
  # filter(!is.na(juv.age),
  #        !is.na(year),
  #        year > 2000) %>%
  # slice_sample(n = 5000) %>%
  mutate(yday.may1 =  if_else(is.na(date), 1, yday(date) - yday("2025-05-01")+1), 
         doy.dec = if_else(yday.may1 > 0, yday.may1, yday.may1+365)/365,
         tot.age.2 = if_else(yday.may1 > 0, tot.age+doy.dec, tot.age+doy.dec-1),
  #       ) %>% # no 0 year olds caught before May 1
  # filter(tot.age == 0, 
  #        yday.may1 <0)
         smo.age = if_else(age.type == "both", juv.age, NA),
         sea.age = if_else(is.na(sea.age), 0, sea.age),
         #age.index = as.integer(tot.age + 1), # + 1 to index age 0 individuals
         age.index = trunc(tot.age.2)+1, # + 1 to index age 0 individuals
         hatch.year = year-tot.age,
         hatch.year.f = as.numeric(factor(hatch.year, levels = sort(unique(hatch.year)), labels = seq_along(sort(unique(hatch.year))))),
         year.f = as.integer(factor(year)),
         spat.unit = as.integer(factor(spat.unit))) %>%
  select(smo.age, tot.age, length_mm, age.index, year, spat.unit, hatch.year, hatch.year.f, doy.dec, tot.age.2, date) %>%
  mutate(ind.id = row_number()) # ind id

n.ind <- nrow(data) # n individuals
npars <- 2 # n parameters in the LKJ dmnorm 
nsu <- length(unique(data$spat.unit))
#age.index = data$age.index
age.index = data$age.index
n.ages = length(unique(data$age.index))+1
nK = length(unique(data$smo.age %>% na.omit()))
# get max year by, needed as hatch.year is integer and the last year is not (simpler solution?).
maxyear <- max(data$hatch.year.f) + data %>%
  filter(year == max(year)) %>%
  filter(tot.age == max(tot.age)) %>%
  distinct(tot.age) %>% pull(tot.age)
minyear = min(data$hatch.year.f)

inits <- function(){
  list(#l.mu = matrix(abs(rnorm(n.ages * n.ind,500,25)), nrow = n.ages, ncol = n.ind), # this is not correct, only inits of =<tot.age
       sig.l = abs(rnorm(1,250,25)),
       sig.par = c(rlnorm(1,0,1),
                   rlnorm(1,0,1)),
       mu.par = c(rnorm(1,7,0.2),
                  rnorm(1,-1,0.6)),
       lb.mu = abs(rnorm(nsu,20,5)),
       g = matrix(rnorm(nsu * length(minyear:maxyear),4,.2), nrow = nsu, ncol = length(minyear:maxyear))
      # smo.age = round(runif(n.ind,0,6))
       #Ustar, #U, #par
       )
  }
       
# build model
biph5.con.model <- nimbleModel(biph5.con.code,
                           constants = list(npars = npars,
                                            nobs = nrow(data),
                                            n.ind = n.ind, # change when >1 sample per ind.id
                                            nK = nK,
                                            alpha = c(1, 2, 3, 2, 1, 1)[1:nK],
                                            nsu = nsu,
                                            #tot.age = data$tot.age,
                                            tot.age.2 = data$tot.age.2,
                                            age.index = age.index,
                                            doy.dec = data$doy.dec,
                                            ind.id = data$ind.id,
                                            hatch.year.f = data$hatch.year.f,
                                            minyear = minyear,
                                            maxyear = maxyear,
                                            spat.unit = data$spat.unit),
                           inits = inits(),
                           data = data %>% select(-ind.id,-age.index,-spat.unit,-hatch.year, -year, -tot.age,-tot.age.2, -hatch.year.f, -doy.dec, -date))

# data %>%
#   distinct(tot.age, tot.age.2, age.index, doy.dec, date) %>%
#   slice_sample(n = 100)

biph5.con.model$simulate("sig.l")
biph5.con.model$simulate("mu.par")
biph5.con.model$simulate("sig.par")
biph5.con.model$simulate("Ustar")
biph5.con.model$simulate("U")
biph5.con.model$simulate("smp")
biph5.con.model$simulate("smo.age")
biph5.con.model$simulate("lb.mu")
biph5.con.model$simulate("g")
biph5.con.model$simulate("par")
biph5.con.model$simulate("l.mu")

# biph5.con.model$calculate("sig.l")
# biph5.con.model$calculate("mu.par[1]")
# biph5.con.model$calculate("sig.par")
# biph5.con.model$calculate("Ustar")
# biph5.con.model$calculate("U")
# biph5.con.model$calculate("smp")
# biph5.con.model$calculate("smo.age")
# biph5.con.model$calculate("lb.mu")
# biph5.con.model$calculate("g")
# biph5.con.model$calculate("par")
# biph5.con.model$calculate("l.mu")

biph5.con.model$sig.l
biph5.con.model$U
biph5.con.model$Ustar
biph5.con.model$mu.par
biph5.con.model$sig.par
biph5.con.model$smp
biph5.con.model$smo.age
#data$smo.age[1:20]
biph5.con.model$lb.mu
exp(biph5.con.model$g)
exp(biph5.con.model$par[1:11,1:25,1])
exp(biph5.con.model$par[1:11,1:20,2])
# Negative l.mus
sum(is.na(biph5.con.model$l.mu))/length(biph5.con.model$l.mu) # percent negatives
# NAs are after
biph5.con.model$l.mu[1,1000:1100] 
max(biph5.con.model$l.mu, na.rm = TRUE) 

#data %>%
#  filter(ind.id %in% c(1994))
```
### Sample g, par, lb.mu, sig.l

```{r samp b1}
# compile model
biph5.con.c <- compileNimble(biph5.con.model)

# configure..
biph5.con.confmcmc <- configureMCMC(biph5.con.c, monitors = c("g","par","lb.mu","sig.l"), project = biph5.con.model, useConjugacy = FALSE, enableWAIC = TRUE)

 # and build MCMC
biph5.con.mcmc <- buildMCMC(biph5.con.confmcmc, project = biph5.con.model)

# compile mcmc
biph5.con.mcmcc <- compileNimble(biph5.con.mcmc, project = biph5.con.model)

# MCMC Samples
biph5.con.samp <- runMCMC(biph5.con.mcmcc, niter = 6000, nburnin = 3000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)

```

### Sample l.mu
```{r samp b2}
# configure..
biph5.con.confmcmc <- configureMCMC(biph5.con.model, monitors = c("l.mu"), project = biph5.con.model, useConjugacy = FALSE, enableWAIC = TRUE)

# and build MCMC
biph5.con.mcmc.lmu <- buildMCMC(biph5.con.confmcmc, project = "biph5.con.model", project = biph5.con.model)

# compile mcmc (and specify the project model)
biph5.con.mcmcc.lmu <- compileNimble(biph5.con.mcmc.lmu, project = biph5.con.model, resetFunctions = TRUE)

# MCMC Samples
biph5.con.samp.lmu <- runMCMC(biph5.con.mcmcc.lmu, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)

```

### Sample smo.age
```{r samp 3b}
# configure
biph5.con.confmcmc <- configureMCMC(biph5.con.model, monitors = c("smo.age"), useConjugacy = FALSE, enableWAIC = TRUE)

# biph5.con.confmcmc <- configureMCMC(biph5.con.c, monitors = c("smo.age"), useConjugacy = FALSE, project = biph5.con.model, enableWAIC = TRUE)

# and build MCMC
biph5.con.mcmc.smo <- buildMCMC(biph5.con.confmcmc)

# compile mcmc (and specify the project model)
biph5.con.mcmcc.smo <- compileNimble(biph5.con.mcmc.smo)

# MCMC Samples
#biph5.samp.lmu <- runMCMC(biph5.mcmcc.lmu, niter = 10000, nburnin = 5000, nchains = 2, samplesAsCodaMCMC = TRUE)
biph5.con.samp.smo <- runMCMC(biph5.con.mcmcc.smo, niter = 6000, nburnin = 3000, nchains = 2, samplesAsCodaMCMC = TRUE, WAIC = TRUE)
```

### Check g, par, lb.mu, sig.l
```{r check bpars}
samp <- biph5.con.samp$samples
sum <- summary(samp)

# summary of values of sampled parameters
vars_subset <- grep("^g\\[", rownames(sum$statistics), value = TRUE)
summary(exp(sum$statistics[which(rownames(sum$statistics)%in%vars_subset),][,1]))
vars_subset <- grep("^par\\[", rownames(sum$statistics), value = TRUE)
vars_subset1 <- grep("1]", vars_subset, value = TRUE)
summary(exp(sum$statistics[which(rownames(sum$statistics)%in%vars_subset1),][,1]))
vars_subset2 <- grep("2]", vars_subset, value = TRUE)
summary(exp(sum$statistics[which(rownames(sum$statistics)%in%vars_subset2),][,1]))
vars_subset <- grep("^lb.mu\\[", rownames(sum$statistics), value = TRUE)
summary(sum$statistics[which(rownames(sum$statistics)%in%vars_subset),][,1])
vars_subset <- grep("sig.l", rownames(sum$statistics), value = TRUE)
sum$statistics[which(rownames(sum$statistics)%in%vars_subset),]

# some trace plots
rn <- sample(1:500, 10)
vars_subset <- grep("^g\\[", varnames(samp), value = TRUE)[rn]
traceplot(samp[, vars_subset, drop = FALSE])

vars_subset <- grep("^par\\[", varnames(samp), value = TRUE)[rn]
traceplot(samp[, vars_subset, drop = FALSE])

vars_subset <- grep("^lb.mu\\[", varnames(samp), value = TRUE)
traceplot(samp[, vars_subset, drop = FALSE])

vars_subset <- grep("sig.l", varnames(samp), value = TRUE)
traceplot(samp[, vars_subset, drop = FALSE])

```

### Plot g, par...

```{r plot bpars}
# g[spat.unit, hatch.year]
samp %>%
  gather_draws(g[su, hy], sep = ",") %>%
  filter(hy == c(1:5)) %>%
  ggplot() +
  geom_density(aes(x = exp(.value), color = factor(hy))) + # age parameter
  facet_wrap(~su, scales = "free") +
  theme_light()

# lb.mu[spat.unit]
samp %>%
  gather_draws(lb.mu[su], sep = ",") %>%
  #filter(hy == c(1:5)) %>%
  ggplot() +
  geom_density(aes(x = .value, color = factor(su))) + # age parameter
  #facet_wrap(~su, scales = "free") +
  theme_light() +
  labs(title = "lb.mu")

# par 1
samp %>%
  gather_draws(par[su, hy, parm], sep = ",") %>%
  filter(hy == c(1:5),
         parm == 1) %>%
  ggplot() +
  geom_density(aes(x = exp(.value), color = factor(hy))) + # age parameter
  facet_wrap(~su, scales = "free") +
  theme_light() +
  labs("linf")

# par 2
samp %>%
  gather_draws(par[su, hy, parm], sep = ",") %>%
  filter(hy == c(1:5),
         parm == 2) %>%
  ggplot() +
  geom_density(aes(x = .value, color = factor(hy))) + # age parameter
  facet_wrap(~su, scales = "free") +
  theme_light() +
  labs("k")

```

### Check l.mu
```{r check blmu}
#samp.lmu <- biph5.con.samp.lmu
# remove NA nodes (the older-than-catch-age nodes)
cm <- lapply(biph5.con.samp.lmu$samples, as.matrix)
nonNA_nodes <- cm %>% 
  map(~ colnames(.x)[colSums(is.na(.x)) == 0]) %>%
  reduce(intersect) # applies function intersect over elements of the lists returned by map(), i.e. removing the colnames from both lists.
cm_cleaned <- lapply(cm, function(x) x[, nonNA_nodes, drop = FALSE])
rm(cm) # big one
samp.cl <- mcmc.list(lapply(cm_cleaned, mcmc))  

#sum <- summary(samp.cl) # not working when l.mu == NA

# some traces of l.mu
vars_subset <- grep("^l.mu\\[", varnames(samp.cl), value = TRUE)[rn]
traceplot(samp.cl[, vars_subset, drop = FALSE])

```

### Plot predicted vs observed l.mu

```{r b predobs}
# plot observed against predicted
obs <- data %>%
  select(length_mm,tot.age,ind.id,spat.unit) %>%
  mutate(tot.age = round(tot.age),
         observed = length_mm)
gc()
predi <- samp.cl[,1:3000] %>%
  gather_draws(l.mu[age.i,ind.id], sep = ",") %>%
  median_qi() %>%
  rename(predicted = .value) %>%
  mutate(tot.age = age.i-1) %>%
  filter(tot.age %in% obs$tot.age & ind.id %in% obs$ind.id)

plusg.po <- predi %>%
  left_join(obs, by = join_by(tot.age, ind.id)) %>%
  drop_na(observed) %>%
  ggplot(aes(x = observed, y = predicted, color = spat.unit)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_errorbar(aes(ymin = .lower, ymax = .upper), width = 0.2, alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_minimal() +
  ylim(0,1300) +
  xlim(0,1300) 

predi %>%
  left_join(obs, by = join_by(tot.age, ind.id)) %>%
  group_by(ind.id) %>%
  mutate(spat.unit = ifelse(is.na(spat.unit), first(na.omit(spat.unit)), spat.unit)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(tot.age, predicted, color = ind.id, group = ind.id)) +
  geom_point(aes(tot.age, observed, color = ind.id, group = ind.id), alpha = 0.2) +
  facet_wrap(~spat.unit)

```

### Plot smolt ages
```{r }
data %>%
  filter(!is.na(smo.age)) %>%
  select(smo.age) %>%
  mutate(t= "obs") %>%
  bind_rows( biph5.con.samp.smo %>% gather_draws(smo.age[i]) %>% rename(smo.age = .value) %>% mutate(t="pred") ) %>%
  ggplot() +
  geom_density(aes(smo.age, fill = t))
```

## AIC comparison

```{r}
biph5.samp$WAIC$WAIC
biph5.con.samp$WAIC$WAIC
biph5.samp.lmu$WAIC$WAIC
biph5.con.samp.lmu$WAIC$WAIC
biph5.samp.smo$WAIC$WAIC
biph5.con.samp.smo$WAIC$WAIC
```
```{r}
plusg.po + inte.po
```


