---
title: "Salmon Biphasic Growth and Fecundity  model"
author: "Viktor Thunell" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    page-layout: full
    embed-resources: true
    toc: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

## Load libraries

```{r libs}
#| message: false
#| warning: false
#| cache: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog", "devtools","viridis","nls.multstart", "broom", "patchwork", "coda", "boot", "tidybayes","bayesplot", "nimble", "here", "purrr")

# remotes::install_github("nimble-dev/nimble", ref = "devel", subdir = "packages/nimble")

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){
    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  }

invisible(lapply(pkgs, library, character.only = T))

options(ggplot2.continuous.colour = "viridis")
theme_set(theme_minimal()) # check why global theme option not working when rendering

# Set path
color_scheme_set("viridis")
home <- here::here()
```

## Read and prep model data

```{r}
#| message: false
#| warning: false
#| cache: false
#set.seed(123)
## Read data
sallaa <- readRDS(file = paste0(home,"/data/data-for-2-2/salmon-laa_2025-12-21.RData")) %>%
  rename(year = fi_year,
         site = sai_location,
         sea.age = sea_age_year,
         juv.age = juvenile_age_year,
         tot.age = tot_age_year,
         lat = fisa_y_4326,
         lon = fisa_x_4326) %>%
  #filter out tot.age = 0 individuals
  filter(!(sea.age == 0 & juv.age == 0)) %>%
  #filter out non-aged individuals (~90000 individuals)
  filter(!(is.na(sea.age) & is.na(juv.age))) %>%
  # removes smolt.age = 0:
  filter(!(age.type == "both" & juv.age == 0)) %>%
  mutate(hatch.year = year-tot.age) %>% # to filter data based o hatch year instead of year to predict growth pars before catch year
  filter(#!is.na(juv.age), # removes age.type = "sea.only"
         !is.na(year),
         hatch.year > 1997)
  
data.biph <- sallaa %>%
  filter(!(is.na(sex) & age.type == "both")) %>% # removes 27 000 rows
  mutate(yday.may1 = if_else(is.na(date), 1, yday(date) - yday("2025-05-01")+1), 
         doy.dec = if_else(yday.may1 > 0, yday.may1, yday.may1+365)/365,
         tot.age.dec = if_else(yday.may1 > 0, tot.age+doy.dec, tot.age+doy.dec-1),
         smo.age = if_else(age.type == "both", juv.age, NA),
         sea.age = if_else(is.na(sea.age), 0, sea.age),
         g.year = trunc(tot.age.dec)+1, # + 1 to index age 0 individuals
         hatch.year.f = as.numeric(factor(hatch.year, levels = sort(unique(hatch.year)), labels = seq_along(sort(unique(hatch.year))))),
         year.f = as.integer(factor(year)),
         sex = as.integer(if_else(sex == "f", 1,2)),
         su = as.integer(factor(spat.unit))) %>%
  mutate(ind.id = row_number())

data.b <- data.biph %>%
  filter(!is.na(smo.age))

data.j <- data.biph %>%
  filter(age.type == "juve.only") %>%
  rename(length_mm_j = length_mm)


# Fecundity 
salfec <- readRDS(file = paste0(home,"/data/data-for-2-2/salmon-fec_2025-12-17.RData"))

data.fec <- salfec %>%
  #filter(!sai_cou_code == "IE") %>%
  drop_na(spat.unit) %>%
  mutate(n.eggs = if_else(is.na(n.eggs.tot), n.eggs.str, n.eggs.tot),
         su = as.integer(factor(spat.unit)),
         length.lc = log(length_mm) - log(mean(length_mm)))
```

## Length at age 

### Plot length at age data
```{r}
sallaa %>%
  drop_na(sex) %>%
  ggplot(aes(tot.age, length_mm, color = sex)) +
  geom_point() +
  facet_grid(age.type~sex) +
  scale_x_continuous(breaks = seq(0, 13, 1) ) +
  theme_light() 

# If dropping all inds where both sea and sm age is NA
sallaa %>% 
  ggplot() +
  geom_density(aes(x = tot.age), fill = "deeppink", alpha = 0.5) +
  xlim(0, 13) +

sallaa %>% 
  ggplot() + 
  geom_density(aes(x = length_mm), fill = "deeppink", alpha = 0.5)

sallaa %>% 
  drop_na(sea.age,sex) %>%
  filter(spat.unit == "Ireland_west") %>%
  ggplot() +
  geom_density(aes(x = length_mm, fill = factor(sex)), alpha = 0.5) +
  facet_wrap(~tot.age)

# age at smoltification by spatial unit
sallaa %>%
  filter(age.type == "both") %>%
  ggplot() +
  geom_density(aes(x = juv.age, color = spat.unit),  alpha = 0.5) +
  facet_wrap(~spat.unit) +
  xlim(-1, 5)

# juvenile size at age
sallaa %>%
  filter(age.type == "juve.only") %>%
  ggplot(aes(x = juv.age, y =length_mm, color = spat.unit)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm) +
  facet_wrap(~spat.unit)
  
sallaa %>%
  filter(age.type == "both") %>%
  ggplot(aes(tot.age, length_mm, color = spat.unit)) +
  geom_point() +
  facet_wrap(~spat.unit) +
  scale_x_continuous(breaks = seq(0, 13, 1) ) +
  theme_minimal()

sallaa %>%
  filter(age.type == "both",
         !is.na(sex)) %>%
  summarise(ml = mean(length_mm), .by = c(sex,spat.unit,)) %>%
  ggplot() +
  geom_point(aes(x = ml, y = spat.unit, color = factor(sex)),
               outliers = FALSE) +
  theme_minimal() 
  
```

### Define data
```{r}

```


### Source models / Load samples
```{r}

# readRDS("/Users/vitl0001/gitProjects/diasp-lht/data/bipj06_samples_2025-12-20.RData")$WAIC
# readRDS("/Users/vitl0001/gitProjects/diasp-lht/data/bipj06_samples_2025-12-22.RData")$WAIC

laa.samples <- readRDS("/Users/vitl0001/gitProjects/diasp-lht/data/bipj06_samples_2025-12-22.RData")$samples

# # Or source model code and mcmc (load libs and data)
# t <- Sys.time()
# source(file = paste0(home,"/R/model devel/2-2_model_salmon_biphasic_v6.R"))
# Sys.time() - t # approx 12 hours with 12 000 NUTS mcmc samples

```

### Traces and ‘potential scale reduction factor’

```{r}
node.sub <- grep("sig.l", colnames(laa.samples[[1]]), value = TRUE)
laa.samples[, node.sub[], drop = FALSE] %>% mcmc_trace()
laa.samples[, node.sub[], drop = FALSE] %>% gelman.diag()
laa.samples[, node.sub[], drop = FALSE] %>%
effectiveSize()

node.sub <- grep("^g\\[", colnames(laa.samples[[1]]), value = TRUE)
laa.samples[, node.sub[], drop = FALSE] %>% mcmc_trace()
laa.samples[, node.sub[], drop = FALSE] %>% gelman.diag()
print("effective sample size")
laa.samples[, node.sub[], drop = FALSE] %>% effectiveSize()
  
node.sub <- grep("^k.year\\[", colnames(laa.samples[[1]]), value = TRUE)
laa.samples[, node.sub[sample(1:length(node.sub), 36)], drop = FALSE] %>% mcmc_trace()
pl <- laa.samples[, node.sub[], drop = FALSE] %>% gelman.diag()
pl$psrf[  which(pl$psrf[, 1] > 1.1),] 
pl <- laa.samples[, node.sub[], drop = FALSE] %>%
effectiveSize()
print("effective sample size")
pl[which(pl < 500)]

node.sub <- grep("^linf\\[", colnames(laa.samples[[1]]), value = TRUE)
laa.samples[, node.sub[], drop = FALSE] %>% mcmc_trace()
laa.samples[, node.sub[], drop = FALSE] %>% gelman.diag()
print("effective sample size")
laa.samples[, node.sub[], drop = FALSE] %>% effectiveSize()

node.sub <- grep("^lb.mu\\[", colnames(laa.samples[[1]]), value = TRUE)
laa.samples[, node.sub[], drop = FALSE] %>% mcmc_trace()
laa.samples[, node.sub[], drop = FALSE] %>% gelman.diag()
print("effective sample size")
laa.samples[, node.sub[], drop = FALSE] %>% effectiveSize()

#node.sub <- grep("^Ustar\\[", colnames(laa.samples[[1]]), value = TRUE) 
node.sub <- grep("^R\\[", colnames(laa.samples[[1]]), value = TRUE) 
laa.samples[, node.sub[sample(1:length(node.sub), 36)], drop = FALSE] %>% mcmc_trace()
pl <- laa.samples[, node.sub[], drop = FALSE] %>% gelman.diag(multivariate = FALSE)
pl$psrf[  which(pl$psrf[, 1] > 1.1),] 
pl <- laa.samples[, node.sub[], drop = FALSE] %>%
effectiveSize()
print("effective sample size")
pl[which(pl < 500 & pl > 0)]
```

### Predicted / observed

```{r}
# juveniles
dfj <- laa.samples %>%
  spread_draws(l.mu.j[i], sep = ",") %>%
  median_qi() %>%
  rename(pred.l = l.mu.j,
         ind.id = i) %>%
  ungroup() %>%
  left_join(data.j %>%
              select(length_mm_j,spat.unit) %>%
              rename(obs.l = length_mm_j) %>%
              mutate(ind.id = row_number()), by = "ind.id")

dfj %>%
  ggplot() +
  geom_abline(slope = 1) +
  geom_point(aes(x = obs.l, y = pred.l, color = spat.unit),shape = 4,alpha = 0.5) +
  theme_minimal() +
  facet_wrap(~spat.unit)

# l.mu[su[i],smo.age[i],coh[i],g.year[i]]
# adults, l.mu[su[i],smo.age[i],coh[i],sex[i],g.year[i]]
dfl <- laa.samples %>%
  spread_draws(l.mu[j,k,l,m,n], sep = ",") %>%
  median_qi() %>%  
  rename(pred.l = l.mu,
         su = j,
         smo.age = k,
         hatch.year.f = l,
         sex = m,
         g.year = n) %>%
  ungroup() %>%
  left_join(data.biph %>%
              select(length_mm,spat.unit,smo.age,hatch.year.f,sex,g.year,su) %>%
              rename(obs.l = length_mm)) 

dfl %>%
  ggplot() +
  geom_abline(slope = 1) +
  geom_point(aes(x = obs.l, y = pred.l, color = spat.unit),shape = 4,alpha = 0.5) +
  theme_minimal() +
  facet_wrap(~spat.unit)

```

### length at age variables

```{r}
# l.mu[j,k,l,m,n] <- l.mu[j,k,l,m,n-1] + step(k - n)*exp(g[j]) +
#             (1-step(k - n)) * ((exp(linf[j,m]) - l.mu[j,k,l,m,n-1])*(1 - 
# exp(-exp(k.year[j,m,(l+(n-1))]))))
  
# length_mm_j[ii] ~ dnorm(l.mu.j[ii], sd = sig.lj)
# l.mu.j[ii] <- lb.mu[suj[ii]] + exp(g[suj[ii]])*gj.year[ii]

laa.samples %>%
  gather_draws(linf[su,sex], sep = ",") %>%
  ungroup() %>%
  left_join(data.biph %>% distinct(su,spat.unit)) %>%
  ggplot() +
  geom_density(aes(x = exp(.value), color = factor(sex))) +
  facet_wrap(~spat.unit, scales = "free", ncol = 3) +
  theme_minimal() 

laa.samples %>%
  gather_draws(g[su]) %>%
  ungroup() %>%
  left_join(data.biph %>% distinct(su,spat.unit)) %>%
  ggplot() +
  geom_density(aes(x = exp(.value), color = spat.unit)) +
  theme_minimal() 

laa.samples %>%
  gather_draws(sig.l,sig.lj) %>%
  ggplot() +
  geom_density(aes(x = .value, color = .variable)) +
  facet_wrap(~.variable, scales = "free", ncol = 3) +
  theme_minimal() 

laa.samples %>%
  gather_draws(k.year[su,sex,year], sep = ",") %>%
  ungroup() %>%
  left_join(data.biph %>% distinct(su,spat.unit)) %>%
  mutate(sex = ifelse(sex == 1, "m","f")) %>%
  ggplot() +
  geom_boxplot(aes(x = exp(.value), y = spat.unit, color = factor(sex)),
               outliers = FALSE) +
  theme_minimal() +
  #xlim(0,)+
  coord_cartesian(xlim = c(-0.5,4)) +
  labs(y = "", x = "k")

laa.samples %>%
  gather_draws(linf[su,sex], sep = ",") %>%
  ungroup() %>%
  left_join(data.biph %>% distinct(su,spat.unit)) %>%
  mutate(sex = ifelse(sex == 1, "m","f")) %>%
  ggplot() +
  geom_boxplot(aes(x = exp(.value), y = spat.unit, color = factor(sex)),
               outliers = FALSE) +
  theme_minimal() +
  labs(y = "", x = "linf")

```
### k over time

```{r}
tmpK <- laa.samples %>%
  gather_draws(k.year[su, sex, gyr], sep = ",") %>%
  left_join(data.b %>% distinct(su, spat.unit)) %>%
  mutate(sex = ifelse(sex == 1, "m","f")) 

tmpK %>%
  filter(gyr < 28) %>%
  ggplot() +
  geom_boxplot(aes(x = factor(gyr), y = exp(.value), color = factor(sex)),
               outliers = FALSE) +
  #scale_x_discrete(labels = hy.labs,breaks = function(x) x[as.numeric(x) %% 2 == 0] ) +
  facet_wrap(~sex) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90,hjust = 1, size = rel(.75))) +
  labs(x = "year", y = "k")

tmpK %>%
  filter(#sex == 2,
         gyr < 28) %>%
  ggplot() +
  geom_boxplot(aes(x = factor(gyr), y = exp(.value), color = factor(sex)), alpha = 0.5,
               outliers = FALSE) +
  #scale_x_discrete(labels = hy.labs,breaks = function(x) x[as.numeric(x) %% 2 == 0] ) +
  facet_wrap(~spat.unit, scales = "free") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90,hjust = 1, size = rel(.75))) +
  labs(x = "year", y = "k", title = "males")
  
tmpK %>%
  filter(sex == "f",
         gyr < 28) %>%
  ggplot() +
  geom_boxplot(aes(x = factor(gyr), y = exp(.value)),
               outliers = FALSE) +
  #scale_x_discrete(labels = hy.labs,breaks = function(x) x[as.numeric(x) %% 2 == 0] ) +
  facet_wrap(~spat.unit, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90,hjust = 1, size = rel(.75))) +
  labs(x = "year", y = "k", title = "females")


```

## Fecundity 

### Plot fecundity data
```{r}
salfec %>%
  ggplot() +
  geom_point(aes(x = length_mm, y = n.eggs.tot), alpha = 0.3, shape = 4) +
  geom_point(aes(x = length_mm, y = n.eggs.str), color = "deeppink", alpha = 0.2) +
  facet_wrap(~spat.unit)

salfec %>%
  ggplot() +
  geom_density(aes(x = n.eggs.tot), fill = "grey", alpha = 0.5) +
  geom_density(aes(x = n.eggs.str), fill = "deeppink", alpha = 0.5) +
  facet_wrap(~spat.unit)

salfec %>%
  mutate(n.eggs = if_else(is.na(n.eggs.tot), n.eggs.str, n.eggs.tot)) %>%
  ggplot() +
  geom_point(aes(x = log(length_mm)-log(mean(length_mm)), y = log(n.eggs)), alpha = 0.5) +
  facet_wrap(~spat.unit)

salfec %>%
  ggplot() +
  geom_density(aes(x = length_mm), fill = "deeppink", alpha = 0.5) +
  facet_wrap(~spat.unit) +
  xlim(200,NA) +
  theme_minimal()
```

### Source models / Load samples
```{r}
readRDS("/Users/vitl0001/gitProjects/diasp-lht/data/fec_samples_1219.RData")$WAIC
readRDS("/Users/vitl0001/gitProjects/diasp-lht/data/fec_samples_1217.RData")$WAIC
fec.samples <- readRDS("/Users/vitl0001/gitProjects/diasp-lht/data/fec_samples_1219.RData")$samples

fec.samples[which(min(effectiveSize(fec.samples)) < 2000)]

# # Or source model code and mcmc (load libs and data)
# source(file = paste0(home,"/R/model devel/2-2_model_salmon_fecundity_v3.R"))
# t <- Sys.time()
# Sys.time() - t # approx 12 hours with 12 000 NUTS mcmc samples

```

### Traces and ‘potential scale reduction factor’

```{r}
node.sub <- grep("^(a|b|r)", colnames(fec.samples[[1]]), value = TRUE)
fec.samples[, node.sub[], drop = FALSE] %>%
mcmc_trace()
fec.samples[, node.sub[], drop = FALSE] %>%
gelman.diag()
```

### Predicted / observed

```{r}
dff <- fec.samples %>%
  spread_draws(mu[i]) %>%
  median_qi() %>%
  rename(pred.ne = mu,
         ind.id = i) %>%
  ungroup() %>%
  left_join(data.fec %>% 
              select(length_mm,n.eggs,spat.unit) %>% 
              rename(obs.ne = n.eggs) %>%
              mutate(ind.id = row_number()))

dff %>%
  ggplot() +
  geom_abline(slope = 1) +
  geom_point(aes(x = obs.ne, y = pred.ne, color = spat.unit),shape = 4,alpha = 0.5) +
  theme_minimal() +
  facet_wrap(~spat.unit)
```

### Fecundity variables

```{r}
fec.samples %>%
  gather_draws(b[su], a[su]) %>%
  ungroup() %>%
  left_join(data.fec %>% distinct(su,spat.unit)) %>%
  mutate(sdval = sd(.value)) %>%
  ggplot() +
  geom_density(aes(x = .value, color = spat.unit)) +
  facet_wrap(~.variable, scales = "free", ncol = 3) +
  theme_minimal() 
  
fec.samples %>%
  gather_draws(mu.a,mu.b,sd.a,sd.b,r) %>%
  ggplot() +
  geom_density(aes(x = .value, color = .variable)) +
  facet_wrap(~.variable, scales = "free", ncol = 3) +
  theme_minimal() 

meanf <- fec.samples %>%
  gather_draws(a[su], b[su]) %>%
  ungroup() %>%
  summarise(meanf = mean(.value), .by = .variable)

fec.samples %>%
  gather_draws(b[su], a[su]) %>%
  ungroup() %>%
  left_join(data.fec %>% distinct(su,spat.unit)) %>%
  ggplot() +
  geom_boxplot(aes(x = .value, y = spat.unit, color = spat.unit),
               outliers = FALSE, show.legend = FALSE) +
  facet_wrap(~.variable, scales = "free_x") +
  theme_minimal() +
  geom_vline(data = meanf, aes(xintercept = meanf), color = "darkgrey", size = 0.5) +
  theme(axis.text.x = element_text(angle = 90,hjust = 1, size = rel(.75))) +
  labs(y = "", x = "LAS [mm]")
```

