---
title: "Sötebasen salmon data for DIASPARA 2.2"
author: "Viktor Thunell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: false
---

## Load libraries

```{r libs}
#| message: true
#| warning: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog","devtools","viridis","patchwork","stringi","readxl","broom","sdmTMB") 

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){
  
    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  }

invisible(lapply(pkgs, library, character.only = T))

options(ggplot2.continuous.colour = "viridis")
devtools::source_url("https://raw.githubusercontent.com/VThunell/diasp-lht/main/R/functions/map-plot.R") # message of SHA-1 hash for file

# Set path
home <- here::here()
```

## Sötebasen data 

1. Sweden SLU database "Sötebasen"
    + At locations across Sweden on both the Baltic side and in the Western sea 
    + From recreational and commercial catches and scientific surveys
    + We assume these are fish returning to spawn but catches are sometime coastal(!)
    + Coordinates of sai_location (river or sometimes region, i.e. Baltic Sea or West coast). This is not the exact catch place.

```{r}
#| warning: true
#| cache: false
#| output: false

# Sweden SLU database "Sötebasen"
sbas.sallaa <- read_delim(file = "/Users/vitl0001/Documents/Projects/DIASPARA/Incoming data/salmon/LaxindividerSötebasen.csv", delim = ";", locale = locale(encoding="latin1"), col_types = cols(Märkning2 = col_character(), Märke2Nr = col_character(), ÅlderLek3 = col_character())) %>%
  # remove individuals without length or sea age
  rename(sea_age_year = AdultÅlder,
         juvenile_age_year = SmoltÅlder,
         length = Längd,
         weight_g = Vikt,
         year = Årtal,
         origin = Ursprung,
         sai_location = VattenNamn,
         stage = Stadium,
         ) %>%
  mutate(plus.growth = if_else(Pluszon == "Ja", 1, 0),
         sai_cou_code = "SWE",
         date = as.Date(FångstDatum),
         sex = if_else(Kön %in% c("m","f"), Kön, NA),
         origin = case_when(origin == "Odlad" ~ "reared",
                            origin == "Vild" ~ "wild",
                            .default = origin))

```
## sb total sb
```{r}
sbas.sallaa %>%
  mutate(is.aged = if_else((is.na(juvenile_age_year) & is.na(sea_age_year)), 0, 1),
         has.len = if_else(is.na(length), 0, 1),
         has.wei = if_else(is.na(weight_g), 0, 1)) %>%
  summarise(n.tot = n(), 
            n.aged = sum(is.aged), 
            n.leng = sum(has.len), 
            n.weig = sum(has.wei)) 
unique(sbas.sallaa$Serie)
```

```{r}
sbas.sallaa %>%
  #filter((is.na(juvenile_age_year) & is.na(sea_age_year))) %>%
  #filter(!is.na(length)) %>%
  summarise(n = n(), .by= "Syfte") %>%
  arrange(-n)
```

# sb diaspara
```{r}
dia.df <- sbas.sallaa %>%
  filter(!(is.na(juvenile_age_year) & is.na(sea_age_year)),
         !is.na(length),
         !sai_location == c("Vättern","Vänern")) 

dia.df %>%
  summarise(n = n(), .by = Syfte) %>%
  arrange(-n)
```

```{r}
dia.df %>%
  summarise(n = n(), .by = c(Syfte, year)) %>%
  ggplot(aes(year, n, fill = factor(Syfte))) +
  geom_bar(stat="identity", position = "stack")
  

```

## Syfte = "Ingen uppgift"
```{r}
dia.df %>%
  filter(Syfte == "Ingen uppgift") %>%
  distinct(AnstrTyp,InsamlMetod)

dia.df %>%
  filter(Syfte == "Ingen uppgift") %>%
  summarise(n = n(), .by = c(sai_location, year)) %>%
  ggplot(aes(year, n, fill = factor(sai_location))) +
  geom_bar(stat="identity", position = "stack")
```

```{r}
plot_map_Nord +
  geom_point(data = dia.df %>% 
               filter(Syfte == "Ingen uppgift") %>%
               drop_na(WGS84_N_Vatten,WGS84_E_Vatten) %>% 
               add_utm_columns(ll_names = c("WGS84_E_Vatten", "WGS84_N_Vatten"), utm_crs = 32633), 
             aes(X*1000, Y*1000), size = 0.3, color = "#440154FF") +
  labs(title = "Syfte == Ingen uppgift")

```

## Vindelälven
```{r}
sbas.sallaa %>%
  filter(str_detect(Plats, "\\bvindel\\b") | str_detect(sai_location, "\\bvindel\\b"))

```
